Equations & R Code

### 2017-03-09 manuscript/coexistence-stability_SI.Rmd
SI writing with PBA initial comments

---
layout: 11pt
header-includes:
  - \usepackage{lineno}
  - \linenumbers
  - \usepackage{setspace}
  - \usepackage{todonotes}
  - \onehalfspacing
  - \usepackage{rotating}
  - \usepackage{color, soul}
  - \usepackage[font={normalsize},labelfont={bf},labelsep=quad]{caption}
  - \usepackage{tikz}
  - \usepackage{bm,mathrsfs}
  - \usepackage{mathptmx} 
bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
csl: components/ecology_letters.csl

@@

\section{Exploring Parameter Space}
In the main text we presented results that most clearly demonstrated our main point, sacrificing some quantitative rigor in terms of exploring the parameter space of the model.
In many cases, altering model parameters, or making them asymmetric among species, makes coexistence more difficult. For example, as we show in Figure 3 in the main text, imposing competitive hierarchies makes it more difficult for species to coexist, but it does not impact our conclusion that a positive diversity--ecosystem variability relationship is possible.
We imposed competitive hierarchies by making the live-to-dormant biomass fractions unique for each species, but any parameter that controls population growth would do the same.
For example, if we make dormant mortality rates assymetric among species we achieve similar results as shown in Figure 3 (Fig. S1-1).

@@

While the results of competitive assymetries among species are rather intuitive, the sensitivity of our results to changes in the absolute values of parameters is less intuitive and requires exploration.
We took a targeted approach to exploring parameter space by focusing on particular processes.
First, we examined the sensitivity of our results to parameters that control the strength of each coexistence mechanism.
Second, we examined the sensitivity of our results to parameters that previous theory have identified as important for the diversity--ecosystem variability relationship.
We describe each in turn.

\subsection{Parameter sensitivity: coexistence strength}
For the storage effect, coexistence strength declines as:

@@

Our results are qualitatively similar: (1) increasing environmental variation allows more species to coexist, creating a positive relationship between species richness and ecosystem variability; (2) and at any given level of environmental variability, it is always better to have more species, creating a within-site negative relationship between species richness and ecosystem variability (Fig. S1-2).
Although the qualitative patterns are similar, increasing the dormant mortality rate has two interesting effects: (1) it makes coexistence more difficult, (2) it reduces the absolute value of ecosystem *CV* by reducing mean population size a little bit and by reducing the temporal standard deviation a lot, and (3) it weakens the buffering effect of additional species (compare the spread of the lines between top and bottom panels of Fig. S1-2).
The second result occurs because population fluctuations are reduced when very little biomass can be activated to the live stage at the beginning of each season.
That is, in terms of an annual plant, germination of live biomass is always a fraction of a very small number.
The third result is a consequence of the second: populations are not fluctuating that much, so total *CV* can only be reduced by so much with the addition of another species.

@@

For relative nonlinearity, coexistence strength declines as:

@@

As expected, lower mean levels of resource make coexistence more difficult because resource acquisitive species are never favored (Fig. S1-3).
Our general conclusion remains, however, that positive diversity--ecosystem variability relationships are possible.
However, the occurence a of positive diversity--ecosystem variability relationship depends on amount of resource variability being sampled over and the traits of individual species (e.g., resource uptake curves). 
In other words, the patterns we report in the main text require that species coexistence be possible given the environmental conditions.

@@

\subsection{Parameter sensitivity: diversity--ecosystem variability parameters}
\textcolor{red}{THIS SECTION IN PROGRESS...}
Previous theory [@Loreau2013] identifies three main mechanisms by which diversity can reduce ecosystem variability:

@@

*CV* scaled by the mean, and we don't have demographic stochasticity because we are assuming large populations, so reducing overall competition to get over-yielding doesn't matter here...

@@

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[width=4in]{./components/regional_diversity_stability_storage_effect_8species.png}
  \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect. Results are from simulations where environmental variance determines the number species that coexist in a community (e.g., a ``regional'' relationship. Colored points show results from individual simulations and the gray points with connecting line show the mean at each level of environmental variance.}
\end{figure}

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[width=3in]{./components/fourspp_Ruptake_relnonlin.png}
  \caption{Resource uptake curves for each species (represented by different colors) as used in relative nonlinearity simulations. The equation for resource uptake is: $f_{i}(R) = r_{i}R^{a_{i}} / (b_{i}^{a_{i}}+R^{a_{i}})$. Parameter values are as follows. Species 1: \emph{r} = 0.2, \emph{a} = 2, \emph{b} = 2.5; Species 2: \emph{r} = 1, \emph{a} = 5, \emph{b} = 20; Species 3: \emph{r} = 2, \emph{a} = 10, \emph{b} = 30; Species 4: \emph{r} = 5, \emph{a} = 25, \emph{b} = 45.}
\end{figure}

>>>>

---
layout: 11pt
header-includes:
  - \usepackage{lineno}
  - \linenumbers
  - \usepackage{setspace}
  - \usepackage{todonotes}
  - \onehalfspacing
  - \usepackage{rotating}
  - \usepackage{color, soul}
  - \usepackage[font={small},labelfont={bf},labelsep=quad]{caption}
  - \usepackage{tikz}
  - \usepackage{bm,mathrsfs}
  - \usepackage{mathptmx} 
bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
csl: components/ecology_letters.csl

@@

\section{Sensitivity of Results to Parameters}
We relied on numerical simulations to reach the conclusions presented in the main text, which means our results could be sensitive to the particular parameter values chosen.
Here we assess the sensitivity of the results by taking a targeted approach to exploring parameter space.
Specifically, we examine the sensitivity of our results to parameters that control the strength of each coexistence mechanism.

In the main text we did explore some parameter space by conducting simulations across gradients of environmental variance (whether the cue or the resource), correlations of species' responses to environmental conditions, and competitive asymmetries.
In many cases, altering model parameters, or making them asymmetric among species, makes coexistence more difficult. 
For example, as we show in Figure 3 in the main text, imposing competitive hierarchies makes it more difficult for species to coexist, but it does not impact our conclusion that a positive diversity--ecosystem variability relationship is possible.
We imposed competitive hierarchies by making the live-to-dormant biomass fractions unique for each species, but any parameter that controls population growth would do the same.
For example, if we make dormant mortality rates assymetric among species we achieve similar results as shown in Figure 3 (Fig. S1-1).

@@

This result (Fig. S1-x and Figure 3 in the main text) reinforces the notion that our conclusion about a positive diversity--ecosystem variability relationship is contingent upon the regional community consisting of species that *can* coexist at some level of observable environmental variance.
In other words, if there is no species that can take advantage of more variable conditions, then fluctuation-dependent coexistence is not possible and neither are the results we present in the main text.

While the results of competitive assymetries among species are rather intuitive, the sensitivity of our results to changes in the absolute values of parameters is less intuitive and requires exploration.
Below we describe simulations designed to test whether our results change qualitatively or quantitatively due to changes in the absolute values of select parameters.
We chose parameters to investigate based on their influence on species coexistence for each mechanism, assuming our results would be most sensitive to those parameters.

\subsection{Parameter sensitivity: storage effect}
For the storage effect, coexistence strength declines as:

@@

Our results are qualitatively similar: (1) increasing environmental variation allows more species to coexist, creating a positive relationship between species richness and ecosystem variability (richness increases along x-axes within panels); (2) and at any given level of environmental variability, it is always better to have more species, creating a within-site negative relationship between species richness and ecosystem variability (Fig. S1-2).
Although the qualitative patterns are similar, increasing the dormant mortality rate has two interesting effects: (1) it makes coexistence more difficult, (2) it reduces the absolute value of ecosystem *CV* by reducing mean population size a little bit and by reducing the temporal standard deviation a lot, and (3) it weakens the buffering effect of additional species (compare the spread of the lines between top and bottom panels of Fig. S1-2).
The second result occurs because population fluctuations are reduced when very little biomass can be activated to the live stage at the beginning of each season.
That is, in terms of an annual plant, germination of live biomass is always a fraction of a very small number.
The third result is a consequence of the second: populations are not fluctuating that much, so total *CV* can only be reduced by so much with the addition of another species.

@@

\subsection{Parameter sensitivity: relative nonlinearity}
For relative nonlinearity, coexistence strength declines as:

@@

As expected, lower mean levels of resource make coexistence more difficult because resource acquisitive species are never favored (Fig. S1-3).
Our general conclusion remains, however, that positive diversity--ecosystem variability relationships are possible.
However, the occurence a of positive diversity--ecosystem variability relationship depends on amount of resource variability being sampled over and the traits of individual species (e.g., resource uptake curves). 
In other words, the patterns we report in the main text require that fluctuation-dependent species coexistence be possible given the environmental conditions and species' traits.

@@

<!---
\subsection{Parameter sensitivity: diversity--ecosystem variability parameters}
\textcolor{red}{THIS SECTION IN PROGRESS...}
Previous theory [@Loreau2013] identifies three main mechanisms by which diversity can reduce ecosystem variability:

@@

*CV* scaled by the mean, and we don't have demographic stochasticity because we are assuming large populations, so reducing overall competition to get over-yielding doesn't matter here...
--->

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[width=3in]{./components/regional_diversity_stability_storage_effect_8species.png}
  \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect. Results are from simulations where environmental variance determines the number species that coexist in a community (e.g., a ``regional'' relationship. Colored points show results from individual simulations and the gray points with connecting line show the mean at each level of environmental variance.}
\end{figure}

@@

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[width=4in]{./components/fourspp_Ruptake_relnonlin.png}
  \caption{Resource uptake curves for each species (represented by different colors) as used in relative nonlinearity simulations. The equation for resource uptake is: $f_{i}(R) = r_{i}R^{a_{i}} / (b_{i}^{a_{i}}+R^{a_{i}})$. Parameter values are as follows. Species 1: \emph{r} = 0.2, \emph{a} = 2, \emph{b} = 2.5; Species 2: \emph{r} = 1, \emph{a} = 5, \emph{b} = 20; Species 3: \emph{r} = 2, \emph{a} = 10, \emph{b} = 30; Species 4: \emph{r} = 5, \emph{a} = 25, \emph{b} = 45.}
\end{figure}

### 2017-03-09 manuscript/coexistence-stability_SI.tex
SI writing with PBA initial comments

\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\usepackage{todonotes}
\onehalfspacing
\usepackage{rotating}
\usepackage{color, soul}
\usepackage[font={normalsize},labelfont={bf},labelsep=quad]{caption}
\usepackage{tikz}
\usepackage{bm,mathrsfs}
\usepackage{mathptmx}

@@

\section{Exploring Parameter Space}

In the main text we presented results that most clearly demonstrated our
main point, sacrificing some quantitative rigor in terms of exploring
the parameter space of the model. In many cases, altering model
parameters, or making them asymmetric among species, makes coexistence
more difficult. For example, as we show in Figure 3 in the main text,
imposing competitive hierarchies makes it more difficult for species to
coexist, but it does not impact our conclusion that a positive
diversity--ecosystem variability relationship is possible. We imposed
competitive hierarchies by making the live-to-dormant biomass fractions
unique for each species, but any parameter that controls population
growth would do the same. For example, if we make dormant mortality
rates assymetric among species we achieve similar results as shown in
Figure 3 (Fig. S1-1).

@@

While the results of competitive assymetries among species are rather
intuitive, the sensitivity of our results to changes in the absolute
values of parameters is less intuitive and requires exploration. We took
a targeted approach to exploring parameter space by focusing on
particular processes. First, we examined the sensitivity of our results
to parameters that control the strength of each coexistence mechanism.
Second, we examined the sensitivity of our results to parameters that
previous theory have identified as important for the
diversity--ecosystem variability relationship. We describe each in turn.

\subsection{Parameter sensitivity: coexistence strength}

@@

@@

\subsection{Parameter sensitivity: diversity--ecosystem variability parameters}

\textcolor{red}{THIS SECTION IN PROGRESS...} Previous theory (Loreau \&
de Mazancourt 2013) identifies three main mechanisms by which diversity
can reduce ecosystem variability:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Aysnchrony of species' responses to environmental conditions
\item
  Reduced mean competition at the community-level
\item
  Differences in population growth rates, which differentiates the speed
  at which species respond to perturbations.
\end{enumerate}

\emph{CV} scaled by the mean, and we don't have demographic
stochasticity because we are assuming large populations, so reducing
overall competition to get over-yielding doesn't matter here\ldots{}

@@

\hypertarget{refs}{}
\hypertarget{ref-Loreau2013}{}
Loreau, M. \& de Mazancourt, C. (2013). Biodiversity and ecosystem
stability: A synthesis of underlying mechanisms. \emph{Ecology Letters},
16, 106--115.

\hypertarget{ref-Yuan2015}{}
Yuan, C. \& Chesson, P. (2015). The relative importance of relative
nonlinearity and the storage effect in the lottery model.
\emph{Theoretical Population Biology}, 105, 39--52.

>>>>

\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\usepackage{todonotes}
\onehalfspacing
\usepackage{rotating}
\usepackage{color, soul}
\usepackage[font={small},labelfont={bf},labelsep=quad]{caption}
\usepackage{tikz}
\usepackage{bm,mathrsfs}
\usepackage{mathptmx}

@@

\section{Sensitivity of Results to Parameters}

We relied on numerical simulations to reach the conclusions presented in
the main text, which means our results could be sensitive to the
particular parameter values chosen. Here we assess the sensitivity of
the results by taking a targeted approach to exploring parameter space.
Specifically, we examine the sensitivity of our results to parameters
that control the strength of each coexistence mechanism.

In the main text we did explore some parameter space by conducting
simulations across gradients of environmental variance (whether the cue
or the resource), correlations of species' responses to environmental
conditions, and competitive asymmetries. In many cases, altering model
parameters, or making them asymmetric among species, makes coexistence
more difficult. For example, as we show in Figure 3 in the main text,
imposing competitive hierarchies makes it more difficult for species to
coexist, but it does not impact our conclusion that a positive
diversity--ecosystem variability relationship is possible. We imposed
competitive hierarchies by making the live-to-dormant biomass fractions
unique for each species, but any parameter that controls population
growth would do the same. For example, if we make dormant mortality
rates assymetric among species we achieve similar results as shown in
Figure 3 (Fig. S1-1).

@@

This result (Fig. S1-x and Figure 3 in the main text) reinforces the
notion that our conclusion about a positive diversity--ecosystem
variability relationship is contingent upon the regional community
consisting of species that \emph{can} coexist at some level of
observable environmental variance. In other words, if there is no
species that can take advantage of more variable conditions, then
fluctuation-dependent coexistence is not possible and neither are the
results we present in the main text.

While the results of competitive assymetries among species are rather
intuitive, the sensitivity of our results to changes in the absolute
values of parameters is less intuitive and requires exploration. Below
we describe simulations designed to test whether our results change
qualitatively or quantitatively due to changes in the absolute values of
select parameters. We chose parameters to investigate based on their
influence on species coexistence for each mechanism, assuming our
results would be most sensitive to those parameters.

\subsection{Parameter sensitivity: storage effect}

@@

\subsection{Parameter sensitivity: relative nonlinearity}

@@

@@

\hypertarget{refs}{}
\hypertarget{ref-Yuan2015}{}
Yuan, C. \& Chesson, P. (2015). The relative importance of relative
nonlinearity and the storage effect in the lottery model.
\emph{Theoretical Population Biology}, 105, 39--52.
